| Early stopping in clinical PET studies: \n| how to reduce expense and exposure

### 2020-10-22 ArticleDraft/Draft.Rmd
Toned down sentence in recommendation box 2 as per feedback

\begin{tcolorbox}[title=\textbf{Practical recommendations 1},title filled]
{In this simulation set-up we show that it is possible to stop a PET study early if there is a true underlying effect in the population, while keeping the number of false positive findings below 5\%. For a patient-control comparison, we recommend that sequential BF testing not be started until at least 12 subjects per group have been collected, adjusting the BF threshold upwards for larger values of Nmax. Importantly, these settings should only be used when the PET researcher is interested in stopping early when data shows support for H\textsubscript{1}. All BF values showing support in favour of H\textsubscript{0} (BF < 1) should not, with these settings, be interpreted as anything else but inconclusive findings.}
\end{tcolorbox}

@@

\begin{tcolorbox}[title=\textbf{Practical recommendations 2},title filled]
{In this simulation set-up we show that it is possible to stop a PET study early either when there is a true underlying effect in the population, or when the effect is zero. For a patient-control comparison, we recommend that sequential BF testing not start until at least 12 subjects per group have been collected, setting the BF threshold to a minimum of 3, and using a one-sided Cauchy distribution (centered around zero with a \textit{r} = 0.707) as the alternative hypothesis. Using a one-sided test means that the researcher must make an *a priori* prediction of which direction the effect will have, and not change this prediction after the study is started, similar to performing a classical one-sided NHST t-test.}
\end{tcolorbox}

## Objective 3 - Application to real clinical PET data 
The third and final aim of this tutorial is to apply sequential BF testing to a real clinical PET setting. To this purpose, we used already collected data of patients with major depressive disorder and healthy control subjects examined with [$^{11}$C]WAY100635, a radioligand, which binds to the serotonin 1A (5HT1A) receptor [@Parsey2010; @Chen2019]. From this data, we included 40 medication-free patients (mean Age 36.2 (12.9SD); 25 females) and 40 healthy controls (mean Age 37.1 (14.0SD), 25 females), using BP~F~ (specific binding over the free-fraction in plasma) as the primary outcome measure for assessing 5HT1A receptor availability in the brain. The 5HT1A receptor acts inhibitory on serotonergic neurons in the raphe nuclei. A high concentration of receptors in raphe will likely lead to lower transmitter release in serotonergic projection areas. Hence, we hypothesized that patients suffering from MDD show higher 5HT1A receptor availability in the raphe nuclei compared to healthy controls. Here we apply sequential BF testing to examine whether this hypothesis gains enough support to stop data collection before 40 subjects per group are reached.

>>>>

\begin{tcolorbox}[title=\textbf{Practical recommendations 1},title filled]
{In this simulation set-up we show that it is possible to stop a PET study early if there is a true underlying effect in the population, while keeping the number of false positive findings below 5\%. We recommend to not start sequential BF testing until at least 12 subjects per group have been collected, adjusting the BF threshold upwards for larger values of Nmax. Importantly, these settings should only be used when the PET researcher is interested in stopping early when data shows support for H\textsubscript{1}. All BF values showing support in favour of H\textsubscript{0} (BF < 1) should not, with these settings, be interpreted as anything else but inconclusive findings.}
\end{tcolorbox}

@@

\begin{tcolorbox}[title=\textbf{Practical recommendations 2},title filled]
{In this simulation set-up we show that it is possible to stop a PET study early either when there is a true underlying effect in the population, or when the effect is zero. When using the Cauchy distribution to describe the alternative hypothesis, we recommend that sequential BF testing does not start until at least 12 subjects per group have been collected, setting the BF threshold to a minimum of 3, and using a one-sided test. Using a one-sided test means that the researcher must make an *a priori* prediction of which direction the effect will have, and not change this prediction after the study is started, similar to performing a classical one-sided NHST t-test.}
\end{tcolorbox}

## Objective 3 - Application to real clinical PET data 
The third and final aim of this tutorial is to apply sequential BF testing to a real clinical PET setting. To this purpose, we used already collected data of patients with major depressive disorder and healthy control subjects examined with [$^{11}$C]WAY100635, a radioligand which binds to the serotonin 1A (5HT1A) receptor [@Parsey2010; @Chen2019]. From this data, we included 40 medication-free patients (mean Age 36.2 (12.9SD); 25 females) and 40 healthy controls (mean Age 37.1 (14.0SD), 25 females), using BP~F~ (specific binding over the free-fraction in plasma) as the primary outcome measure for assessing 5HT1A receptor availability in the brain. The 5HT1A receptor acts inhibitory on serotonergic neurons in the raphe nuclei. A high concentration of receptors in raphe will likely lead to lower transmitter release in serotonergic projection areas. Hence, we hypothesized that patients suffering from MDD show higher 5HT1A receptor availability in the raphe nuclei compared to healthy controls. Here we apply sequential BF testing to examine whether this hypothesis gains enough support to stop data collection before 40 subjects per group are reached.
